### 2014-02-08 abstract.tex
Polascik's comments

\section*{Abstract}
Prostate cancer (PCa) is the most common non-cutaneous malignancy among men in
the United States and the second leading cause of cancer-related death.
Multi-parametric Magnetic Resonance Imaging (mpMRI) has gained recent
popularity to characterize PCa.  ARFI imaging has the potential to aid in PCa
diagnosis and management by evaluating the structural composition of prostate
zones and tumors based on their stiffness.  In this study, \invivo MR and ARFI
imaging datasets were compared to one another and with gross pathology
measurements made immediately after radical prostatectomy.  Images were
manually segmented in 3D Slicer to delineate the central gland (CG) and
prostate capsule, and 3D models were rendered to evaluate zonal anatomy
dimensions and volumes.  Both imaging modalities showed good correlation
between estimated organ volume and gross pathologic weights.  Ultrasound and MR
total prostate gland volumes were well-correlated (R$^2$ = \MRarfiVolTotalRsq),
but ARFI images yielded prostate volumes that were larger
(\MRarfiVolTotalMeanDiff\% $\pm$ \MRarfiVolTotalStdDiff\%) than MR images,
primarily due to overestimation of the lateral dimension
(\ARFImrTotalLatLatMeanPct $\pm$ \ARFImrTotalLatLatStdPct\%), while differences
in the other dimensions were less significant (\ARFImrTotalAntPostMeanPct $\pm$
\ARFImrTotalAntPostStdPct\%, anterior-to-posterior, and
\ARFImrTotalApexBaseMeanPct $\pm$ \ARFImrTotalApexBaseStdPct\%, apex-to-base).
The ARFI and MR CG volumes were also well-correlated (R$^2$ =
\MRarfiVolCentralRsq).  CG volume differences were attributed to
ARFI underestimation of the apex-to-base axis (\ARFImrCentralApexBaseMeanPct $\pm$
\ARFImrCentralApexBaseStdPct\%), and ARFI overestimation bias in the lateral
dimension (\ARFImrCentralLatLatMeanPct $\pm$ \ARFImrCentralLatLatStdPct\%).
ARFI imaging yielded prostate volumes and dimensions that were well-correlated
with MR T2WI estimates, with biases in the lateral dimension due to poor
contrast caused by peri-prostatic fat.  ARFI imaging is a promising low-cost,
real-time imaging modality that can compliment mpMRI for diagnosis, treatment
planning and management of PCa.

>>>>

\section*{Abstract}
Prostate cancer (PCa) is the most common non-cutaneous malignancy among men in
the United States and the second leading cause of cancer-related death.
Multi-parametric Magnetic Resonance Imaging (mpMRI) has gained recent
popularity to characterize PCa.  ARFI imaging has the potential to aid in PCa
diagnosis and management by evaluating the structural composition of prostate
zones and tumors based on their stiffness.  In this study, \invivo MR and ARFI
imaging datasets were compared to one another and with gross pathology
measurements made immediately after radical prostatectomy.  Images were
manually segmented in 3D Slicer to delineate the central gland (CG) and
prostate capsule, and 3D models were rendered to evaluate zonal anatomy
dimensions and volumes.  Both imaging modalities showed good correlation
between estimated organ volume and gross pathologic weights.  Ultrasound and MR
total prostate gland volumes were well-correlated (R$^2$ = \MRarfiVolTotalRsq),
but ARFI images yielded prostate volumes that were larger
(\MRarfiVolTotalMeanDiff\% $\pm$ \MRarfiVolTotalStdDiff\%) than MR images,
primarily due to overestimation of the lateral dimension
(\ARFImrTotalLatLatMeanPct $\pm$ \ARFImrTotalLatLatStdPct\%), while differences
in the other dimensions were less significant (\ARFImrTotalAntPostMeanPct $\pm$
\ARFImrTotalAntPostStdPct\%, anterior-to-posterior, and
\ARFImrTotalApexBaseMeanPct $\pm$ \ARFImrTotalApexBaseStdPct\%, apex-to-base).
The ARFI and MR CG volumes were also well-correlated (R$^2$ =
\MRarfiVolCentralRsq).  CG volume differences were attributed to ARFI
underestimation of the apex-to-base axis (\ARFImrCentralApexBaseMeanPct $\pm$
\ARFImrCentralApexBaseStdPct\%), and ARFI overestimation bias in the lateral
dimension (\ARFImrCentralLatLatMeanPct $\pm$ \ARFImrCentralLatLatStdPct\%).
ARFI imaging yielded prostate volumes and dimensions that were well-correlated
with MR T2WI estimates, with biases in the lateral dimension due to poor
contrast caused by peri-prostatic fat.  ARFI imaging is a promising low-cost,
portable, real-time imaging modality that can compliment mpMRI for diagnosis,
treatment planning and management of PCa.

### 2014-02-08 intro.tex
Polascik's comments

PCa screening has led to earlier diagnosis of smaller tumors and more localized
disease; however, it is well known that the sensitivity and specificity of PSA
and DRE are not optimal. In addition, DRE has a low predictive value at lower
PSA ranges, and PSA yields many false
positives.~\cite{Gosselaar2007,Gupta2013,Hricak2007} As such, a theoretical
risk of over-diagnosis and treatment of low-grade, and possibly clinically
insignificant, disease exists.  Moreover, due to the random nature of
TRUS-guided systematic biopsies, PCa located outside the routine sampling sites
can be missed and the extent of the cancer might be
underestimated.~\cite{Gupta2013,Cornud2012} For example, in a study by Mufarrij
\etal, 45.9--47.2\% of patients who were candidates for active surveillance,
but underwent radical prostatectomy, had a higher Gleason score on final
histopathology than after TRUS biopsy.~\cite{Mufarrij2010} These inaccuracies
may lead to inappropriate diagnosis, imprecise risk assessment and potentially
avoidable morbidity.

>>>>

PCa screening has led to earlier diagnosis of smaller tumors and more localized
disease; however, it is well known that the sensitivity and specificity of PSA
and DRE are not optimal. In addition, DRE has a low predictive value at lower
PSA ranges, and PSA yields many false
positives.~\cite{Gosselaar2007,Gupta2013,Hricak2007} As such, a theoretical
risk of over-diagnosis and treatment of low-grade, and possibly clinically
insignificant, disease exists.  Moreover, due to the random nature of
TRUS-guided systematic biopsies, PCa located outside the routine sampling sites
can be missed and the extent of the cancer might be
underestimated.~\cite{Gupta2013,Cornud2012} For example, in a study by Mufarrij
\etal, 45.9--47.2\% of patients who were candidates for active surveillance,
but underwent radical prostatectomy, had a higher Gleason score on final
histopathology than after TRUS biopsy.~\cite{Mufarrij2010} These inaccuracies
may lead to missed diagnoses, imprecise risk assessment and potentially
avoidable morbidity.

### 2014-02-08 limitations.tex
Polascik's comments

It should also be noted that all of the prostates in this study contained
varying amounts of PCa, BPH and atrophy, all of which can distort the zonal
anatomy, especially in the case of BPH and central gland morphology.  While
younger, healthier prostates could have been targeted, these healthy organs
would not have been excised for pathology characterization, and the zonal
anatomy of a healthy (young) prostate is expected to be different from the
prostate of a middle-age man, who is the target demographic for PCa screening
imaging and PCa characterization.

>>>>

It should also be noted that all of the prostates in this study contained
varying amounts of PCa, BPH and atrophy, all of which can distort the zonal
anatomy, especially in the case of BPH and central gland morphology.  While
younger, healthier prostates could have been targeted, these healthy organs
would not have been excised for pathology characterization, and the zonal
anatomy of a healthy (young) prostate is expected to be different from the
prostate of a middle-age and older man, who is the target demographic for PCa screening
imaging and PCa characterization.

### 2014-02-08 mr_background.tex
integrtaed Ned's changes
Polascik's comments

\subsection{MR Imaging in Prostate Diagnostics}
MR imaging has been available for use in the workup of patients with PCa since
the early 1980s, but the early work on its diagnostic accuracy is heterogeneous
due to heavy reliance on morphology via T1 and T2-weighted imaging (T2WI). More
recent abilities to include not only anatomic, but also biologic and functional
dynamic parameters, into MR analysis via diffusion-weighted imaging (DWI),
dynamic contrast-enhanced (DCE) imaging or MR spectroscopic imaging (MRSI) are
promising in the future diagnosis and management of PCa.

@@

The prostate can be divided into glandular and non-glandular components. The
glandular components include the peripheral zone (PZ) and the central gland,
which are typically easily distinguishable on T2WI.  Other anatomical markers
such as the urethra, verumontanum and ejaculatory are also often seen on T2WI.

>>>>

\subsection{MR Imaging in Prostate Diagnostics}
MR imaging has been available for use in the workup of patients with PCa since
the early 1980s, but the early work on its diagnostic accuracy is heterogeneous
due to heavy reliance on morphology via T1 and T2-weighted imaging (T2WI). More
recent abilities to include not only anatomic, but also biologic and functional
dynamic parameters, into MR analysis via diffusion-weighted imaging (DWI),
dynamic contrast-enhanced (DCE) imaging or MR spectroscopic imaging (MRSI) are
promising in the future diagnosis and characterization of PCa.

@@

The prostate can be divided into glandular and non-glandular components. The
glandular components include the peripheral zone (PZ) and the central gland,
which are typically easily distinguishable on T2WI.  Other anatomical markers
such as the urethra, verumontanum and ejaculatory ducts are also often seen on
T2WI.
